---
title: "Extract daily point precipitation"
description: >
  This vignette shows how to build a netCDF AWAP data file and then
  extract the point precipitation data at a few locations over a usr defined period.
  The extracted data is then compared with precipitation gauge data.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract daily point precipitation}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
```

```{r setup}
library(AWAPer, warn.conflicts = FALSE)
```

## Make netCDF file
The first step is to create the netCDF files. Here only one netCDF file is created and only for precipitation data (as defined by the following URLS for the source data being set to NA: urlTmin, urlTmax, urlVprp, urlSolarrad) and only between the dates _updateTo_ and _updateTo_. If the latter two dates were not input then data would be downloaded from 1/1/1900 to yesterday. The netCDF file contains grids of daily rainfall for all of Australia and is used below to extract data at points of interest. Often users run _makeNetCDF_file_ once to build netCDF data files that contain all variables over the entire record length (which requires ~5GB disk storage) and then use the netCDFs grids for multiple projects, rather than re-building the netCDF for each project. Also, if _makeNetCDF_file_ is run with the the netCDF file names pointing to existing files and _updateFrom=NA_ then the netCDF files will be updated to yesterday.

```{r}
fnames = makeNetCDF_file(ncdfFilename ='AWAP_demo.nc',
                         updateFrom=as.Date("2010-08-01","%Y-%m-%d"),
                         updateTo=as.Date("2010-10-01","%Y-%m-%d"),
                         urlTmin=NA, urlTmax=NA, urlVprp=NA, urlSolarrad=NA)
```

## Set the points for data extraction
Set coordinates to the location of one groundwater bore and one rainfall gauge and then convert the points to a spatial object and set projection to GDA94.
```{r}
coordinates.data = data.frame( ID =c('Bore-10084446','Rain-63005'),
                               Longitude = c(153.551875, 149.5559),
                               Latitude =  c(-28.517974,-33.4289))

sp::coordinates(coordinates.data) <- ~Longitude + Latitude

sp::proj4string(coordinates.data) = '+proj=longlat +ellps=GRS80 +no_defs'
```

## Extract daily precipitation data 
Extract the daily precipitation at the two locations. The data is extracted from the netCDF file _ncdfFilename_ and between the dates _extractFrom_ and _extractTo_. The other AWAPer variables are not extracted because _getTmin_, _getTmax_, _getVprp_, _getSolarrad_ and _getET_ are set to _F_. Note the netCDF file _ncdfFilename_ must be in the working directory or the full file path must be given.
```{r}
climateData.data = extractCatchmentData(ncdfFilename='AWAP_demo.nc',
                                        extractFrom=as.Date("2010-08-01","%Y-%m-%d"),
                                        extractTo=as.Date("2010-10-01","%Y-%m-%d"),
                                        catchments=coordinates.data,
                                        getTmin=F, getTmax=F, getVprp=F, getSolarrad=F, getET=F)
```

## Plot the daily precipitation at each site
Plot time series of the etracted daily precipitation.

```{r}
for (i in 1:nrow(coordinates.data)){
  filt = climateData.data$CatchmentID.ID == coordinates.data$ID[i]

  data2plot = climateData.data$precip_mm[filt]
  dates = ISOdate(climateData.data$year[filt],climateData.data$month[filt],climateData.data$day[filt])
  
  if (i==1){
    plot(dates, data2plot, main='Extracted precip.', ylab='Precip [mm/d]', xlab='Date [day/month]', type='l',lty = i+1, cex=0.2)
  } else {
    lines(dates, data2plot, lty = i+1)
  }
}
legend('topright',legend=c('Bore-10084446','Rain-63005'),lty=2:3)
```

## Check AWAPer data against precipitation gauge date
Use the following observed precipitation (from gauge_63005) to check the AWAPer data. Note, the observed data is hardcoded in and sourced from the Australian Bureau of Meteorology.

```{r}
obsPrecip <- data.frame(
  year= c(2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010,
          2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010,
          2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010, 2010),
  month = c(8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 9, 9, 9, 9, 9,
            9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 10),
  day = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 1, 2, 3, 4, 5,
          6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 1),
  precip_mm = c(0.6, 5.2, 0.8, 0.4, 0.0, 0.0, 0.0, 0.0, 0.0, 15.8, 15.6, 7.6, 0.7, 0.4, 1.4, 1.0, 0.0, 0.0, 30.4, 1.0, 0.0,
                0.0, 5.0, 2.2, 0.3, 0.8, 13.8, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.5, 13.8, 15.2, 0.4, 0.2, 0.0, 0.0, 12.4, 0.9,
                0.0, 0.0, 0.1, 13.0, 0.2, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.3, 0.0, 0.0, 0.0, 0.0))

# Plot the observed vs AWAPer rainfall at gauge ID 63005
filt2 = climateData.data$CatchmentID.ID=='Rain-63005'
plot(obsPrecip$precip_mm,climateData.data$precip_mm[filt2],
     xlim = c(0,35),ylim = c(0,35),
     main='Obs. vs. AWAPer precip.',
     xlab='Obs. [mm/d]', ylab='AWAPer [mm/d]', cex=0.2)
abline(0,1, col='grey', lty=2)

# Plot the cumulative observed vs AWAPer rainfall at gauge ID 63005
plot(cumsum(obsPrecip$precip_mm),cumsum(climateData.data$precip_mm[filt2]),
     xlim = c(0,175),ylim = c(0,175),
     main='Cum. obs. vs. AWAPer precip.',
     xlab='Obs. [mm]', ylab='AWAPer [mm]', type='l', cex=0.2)
abline(0,1, col='grey', lty=2)
```
